---
title: Barbell Lifts
---

## Introduction
Researchers are attempting to classify different types of one particular action based on the movement of different body parts. Here, an attempt is made to distinguish between several methods of lifting a barbell using machine learning.

## Data
```{r data}
library(caret)
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", "training.csv", "curl")
training <- read.csv("training.csv")
summary(training$classe)
```

The source of the data is http://groupware.les.inf.puc-rio.br/har. 152 variables were tracked over 5 different methods of performing a barbell lift, 
including the acceleration and pitch/roll/yaw of different parts of the arm, a total of 19622 samples. These methods of lifting are:

* exactly according to the specification (Class A)
* throwing the elbows to the front (Class B)
* lifting the dumbbell only halfway (Class C)
* lowering the dumbbell only halfway (Class D)
* throwing the hips to the front (Class E)

## Building the model
```{r train, cache=TRUE}
set.seed(98)
inTrain = createDataPartition(training$classe, p = 0.8)[[1]]
tr = training[inTrain, ]
cv = training[-inTrain, ]
ctrl <- trainControl(method="boot_all")
rf <- train(classe ~ accel_arm_x + accel_arm_y + accel_arm_z + roll_arm + accel_forearm_x +
 accel_forearm_y + accel_forearm_z + accel_dumbbell_x + accel_dumbbell_y + accel_dumbbell_z +
 roll_belt + accel_belt_x + accel_belt_y + accel_belt_z, data=tr,
 method="rf", na.action=na.exclude)
# rf <- train(classe ~ accel_arm_x + accel_arm_y + accel_arm_z + roll_arm + accel_forearm_x + 
#   accel_forearm_y + accel_forearm_z + accel_dumbbell_x + accel_dumbbell_y + accel_dumbbell_z + 
#   roll_belt + accel_belt_x + accel_belt_y + accel_belt_z + var_accel_arm + var_accel_forearm, data=tr, 
#   method="rf", na.action=na.exclude, trControl=ctrl)
rf
```
80% of the given training data was used for training the model, and the remaining 20% was used as cross-validation.
The model built here is a random forest model, using the following variables as predictors for lift type.

 * Arm acceleration in x, y, z directions
 * Forearm acceleration in x, y, z directions
 * Dumbbell acceleration in x, y, z directions
 * Belt acceleration in x, y, z directions
 * Roll of arm, belt

These variables sought to cover all of the monitored parts of the test subjects when they were lifting, and the different directions of acceleration were to address the range of motion associated with each form of barbell lift.

More predictors were attempted, but as the random forest method in R takes much smaller bootstrap samples with additional predictors the accuracy dropped.

## Cross-Validation
```{r cross, cache=TRUE}
pr <- predict(rf, cv)
sprintf("Error = %0.4f", sum(pr != cv$classe)/length(cv$classe))
```
Of the 3923 data points used for cross-validation, the model has an error rate of 2.75%.

## Test Cases
```{r test}
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", "testing.csv", "curl")
testing <- read.csv("testing.csv")
predict(rf, testing)
```
Within the test set of 20 samples, 18 out of 20 were correctly predicted using this model.