---
title: Barbell Lifts
---

## Introduction

## Data
```{r data}
library(caret)
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", "training.csv", "curl")
training <- read.csv("training.csv")
summary(training$classe)
```
<!-- Methods of lifting:
exactly according to the specification (Class A)
throwing the elbows to the front (Class B)
lifting the dumbbell only halfway (Class C)
lowering the dumbbell only halfway (Class D)
throwing the hips to the front (Class E) -->
The source of the data is http://groupware.les.inf.puc-rio.br/har. 152 variables were tracked over 5 different methods of performing a barbell lift, 
including the acceleration and pitch/roll/yaw of different parts of the arm, a total of 19622 samples.

## Building the model
```{r train, cache=TRUE}
inTrain = createDataPartition(training$classe, p = 0.8)[[1]]
tr = training[inTrain, ]
cv = training[-inTrain, ]
rf12 <- train(classe ~ accel_arm_x + accel_arm_y + accel_arm_z + accel_forearm_x + accel_forearm_y + accel_forearm_z + 
    accel_dumbbell_x + accel_dumbbell_y + accel_dumbbell_z + pitch_belt + accel_belt_x + accel_belt_y + accel_belt_z, 
    data=tr, method="rf", na.action=na.exclude)
rf12
```
80% of the given training data was used for training the model, and the remaining 20% was used as cross-validation.
The model built here is a random forest model, using the following variables as predictors for lift type.

 * Arm acceleration in x, y, z directions

 * Forearm acceleration in x, y, z directions

 * Dumbbell acceleration in x, y, z directions

 * Belt acceleration in x, y, z directions

 These variables sought to cover all of the monitored parts of the test subjects when they were lifting, and the different directions of acceleration were to
  address the range of motion associated with each form of barbell lift.

## Cross-Validation
```{r cross, cache=TRUE}
pr <- predict(rf12, cv)
sprintf("Error = %0.4f", sum(pr != cv$classe)/length(cv$classe))
```
Of the 3923 data points used for cross-validation, the model has an error rate of 3.11%.

## Test Cases
```{r test}
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", "testing.csv", "curl")
testing <- read.csv("testing.csv")
predict(rf12, testing)
```
Within the test set of 20 samples, all 20 were correctly predicted using this model.